---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ANTTV

<!-- badges: start -->
<!-- badges: end -->

ANTTV is a suite of R programs to analyse data on New Testament textual variation. It uses the R pipe operator `|>` to chain together functions to produce analysis results.

To get started:

1. Install R on your computer
2. [Optional] To make life easier, install RStudio on your computer
3. Install the ANTTV package
4. Use the programs in the package to analyse the data.

## Installation

You can install the development version of ANTTV from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tjfinney/ANTTV")
```

## Example

To analyse a data frame or distance matrix,

```{r example}
library(ANTTV)
## data frame
read_data_frame("https://zenodo.org/record/6466262/files/Mark.UBS4.csv") |> do_reduction() |> do_dist() |> do_CMDS()
## distance matrix
read_data_frame("https://zenodo.org/record/6505843/files/Acts.UBS2.dist.csv") |> do_CMDS()
```

